---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NASAaccess

<!-- badges: start -->

[![Anaconda-Server Badge](https://anaconda.org/conda-forge/r-nasaaccess/badges/version.svg)](https://anaconda.org/conda-forge/r-nasaaccess) [![Anaconda-Server Badge](https://anaconda.org/conda-forge/r-nasaaccess/badges/downloads.svg)](https://anaconda.org/conda-forge/r-nasaaccess) [![Anaconda-Server Badge](https://anaconda.org/conda-forge/r-nasaaccess/badges/platforms.svg)](https://anaconda.org/conda-forge/r-nasaaccess) [![](https://img.shields.io/readthedocs/nasaaccess?style=social)](https://nasaaccess.readthedocs.io/en/latest/index.html) [![](https://img.shields.io/github/stars/nasa/nasaaccess?style=social)](https://github.com/nasa/NASAaccess)

```{=html}
<!-- ![lifecycle](https://img.shields.io/badge/lifecycle-stable-sucess.svg?style=plastic&logo=appveyor)
<!-- ![GitHub Downloads](https://img.shields.io/github/downloads/nasa/NASAaccess/total?style=plastic)

<!-- badges: end -->
```
[*Ibrahim N. Mohammed*](https://science.gsfc.nasa.gov/sed/bio/ibrahim.mohammed "Ibrahim N. Mohammed")

## **What is NASAaccess?**

*NASAaccess* is a software application in the form of a [R](https://www.r-project.org/about.html) package, a [conda](https://docs.conda.io/en/latest/) package, and a [Tethys](https://www.tethysplatform.org/) web application. *NASAaccess* software can generate gridded ascii tables of climate [CMIP5](https://pcmdi.llnl.gov/mips/cmip5/ "Coupled Model Intercomparison Project Phase 5"), [CMIP6](https://pcmdi.llnl.gov/CMIP6/ "Coupled Model Intercomparison Project Phase 6"), and earth observation remote sensing data ([GPM](https://gpm.nasa.gov/data/directory "Global Precipitation Measurement"), [TRMM](https://gpm.nasa.gov/missions/trmm "Tropical Rainfall Measuring Mission"), [GLDAS](https://ldas.gsfc.nasa.gov/gldas "Global Land Data Assimilation System")) needed to drive various hydrological models (e.g., [SWAT](https://swat.tamu.edu/ "Soil & Water Assessment Tool"), [VIC](https://github.com/UW-Hydro/VIC "Variable Infiltration Capacity"), [RHESSys](https://github.com/RHESSys/RHESSys "The Regional Hydro-Ecological Simulation System"), [DHSVM](https://www.pnnl.gov/projects/distributed-hydrology-soil-vegetation-model "Distributed Hydrology Soil Vegetation Model") ...etc.). The *NASAaccess* Tethys web-based application can be used for accessing, reformatting, and visualizing climate and earth observation remote sensing gridded time series data as well.

## **Where to find the NASAaccess software?**

-   R package can be downloaded from GitHub at <https://github.com/nasa/NASAaccess>.

-   Conda package can be installed directly from *Anaconda* by searching for `r-nasaaccess`.

-   Tethys web-based application can directly installed from GitHub at <https://github.com/imohamme/tethys_nasaaccess>.

## **How NASAaccess software is distributed?**

*NASAaccess* is an open-source software package under [NASA Open Source Agreement v1.3](https://opensource.org/license/nasa1-3-php/).

## **What is needed to install the NASAaccess software on my local machine?**

### **R Library**

On a local machine the user should have installed the following programs as well as setting up a user account. The list below gives a summary of what is needed to be done prior to work with NASAaccess software on any local machine:

-   [Installing R software](https://www.r-project.org/)

-   [Installing Rstudio software](https://posit.co/) (OPTIONAL)

-   *NASAaccess* R package needs a user registration access with [Earthdata](https://www.earthdata.nasa.gov/). Users should set up a registration account(s) with [Earthdata](https://www.earthdata.nasa.gov/) login as well as well as authorizing [NASA](https://www.nasa.gov/ "The National Aeronautics and Space Administration") [GES DISC](https://disc.gsfc.nasa.gov/) data access. Please refer to <https://disc.gsfc.nasa.gov/data-access> for further details.

-   Installing [*curl*](https://curl.se/) software . Since Mac users have [*curl*](https://curl.se/) as part of macOS build, Windows users should make sure that their local machines build have [*curl*](https://curl.se/) installed properly.

-   Checking if you can run [*curl*](https://curl.se/) from your command prompt. Type `curl --help` and you should see the help pages for the [*curl*](https://curl.se/) program once everything is defined correctly.

-   Within Rstudio or R terminal run the following commands to install *NASAaccess*:

    -   `library(devtools)`

    -   `install_github("nasa/NASAaccess", build_vignettes = TRUE)`

    -   `library(NASAaccess)`

Within the Rstudio help tab the user can verify that the package has been installed and browse the help pages of the various functions of *NASAaccess*.

### **Conda Package**

-   After registration with [Earthdata](https://www.earthdata.nasa.gov/) *NASAaccess* software package users should create a reference file (*.netrc*) with [Earthdata](https://www.earthdata.nasa.gov/) credentials stored in it to streamline the retrieval access from [NASA](https://www.nasa.gov/ "The National Aeronautics and Space Administration") servers.

    -   Creating the *.netrc* file at the user machine *Home* directory and storing the user [NASA](https://www.nasa.gov/ "The National Aeronautics and Space Administration") [GES DISC](https://disc.gsfc.nasa.gov/) logging information in it is needed to execute the *NASAaccess* package commands. Accessing data from NASA servers is further explained at [Here](https://wiki.earthdata.nasa.gov/display/EL/How+To+Access+Data+With+cURL+And+Wget).
    
    
-   To install *NASAaccess* package in a conda environment run the following:

``` python
conda install -c conda-forge r-nasaaccess
```

### **Tethys web-based Application**

Full details on installing the web-based application of *NASAaccess* on single machines and local servers can be found at [readthedocs](https://nasaaccess.readthedocs.io/en/latest/nasaaccess_tethys.html).

## **Is there a walk through examples for NASAaccess software?**

Software users are encouraged to visit (<https://imohamme.github.io/NASAaccess/>) to learn more on *NASAaccess* functionality and capabilities.

## **How to cite R NASAaccess package?**

```{r}
citation(package = 'NASAaccess')

```

