---
title: "Introduction"
author: "Ibrahim N. Mohammed"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The _NASAaccess_ package is a simple and needed tool for accessing and processing remote sensing data.
The _NASAaccess_ package has multiple functions that help the user to access and reformat hydrological data for easy ingest into various hydrological models. Since the package functions touch [NASA](https://www.nasa.gov/ "The National Aeronautics and Space Administration") data repositories to retrieve data, the user must set up a registration account with [Earthdata](https://urs.earthdata.nasa.gov/users/new) as well as authorizing [NASA GES DISC](https://disc.gsfc.nasa.gov/data-access) data access. The package user should make sure that his(her) local machines has [curl](https://curl.haxx.se/) installed properly. Further instructions on creating the ".netrc" and ".urs_cookies" files can be accessed at [*How To Access Data NASA data With cURL And Wget wiki page*](https://wiki.earthdata.nasa.gov/display/EL/How+To+Access+Data+With+cURL+And+Wget). Creating the ".netrc" file at the user machine 'Home' directory and storing the [NASA GES DISC](https://disc.gsfc.nasa.gov/) user logging information in it is needed to execute the package commands.

Note: for Windows users the [NASA GES DISC](https://disc.gsfc.nasa.gov/) logging information should be saved in a file named "_netrc" beside the ".netrc" one.


## Index

* [GPM](.\GPM.html)
* [GLDAS](.\GLDAS.html)
* [CMIP5 Climate](.\NEXGDPP.html)

## Built with

```{r}
sessionInfo()

```
