---
title: "Getting started with NEXGDDP-CMIP6 data"
author: "Ibrahim N. Mohammed"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
bibliography: "./bibliography/references.bib"
biblio-style: "apalike"
vignette: >
  %\VignetteIndexEntry{Getting started with NEXGDDP-CMIP6 data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(raster)
library(sf)
library(ggplot2)
#Reading input data
dem_path <- system.file("extdata", "DEM_TX.tif", package = "NASAaccess")
shape_path <- system.file("extdata", "basin.shp", package = "NASAaccess")


dem <- raster(dem_path)
shape <- st_read(shape_path)
```




_NASAaccess_ has a handy tool to access, extract, and reformat climate change data of rainfall and air temperature from [NASA Earth Exchange Global Daily Downscaled Projections NEX-GDDP-CMIP6 AMES servers](https://www.nccs.nasa.gov/services/data-collections/land-based-products/nex-gddp-cmip6) for grids within a specified watershed. 

[NEX-GDDP-CMIP6](https://www.nccs.nasa.gov/sites/default/files/NEX-GDDP-CMIP6-Tech_Note.pdf) dataset is comprised of downscaled climate scenarios for the globe that are derived from the General Circulation Model GCM runs conducted under the Coupled Model Intercomparison Project Phase 6 CMIP6 [@RN1450] and across two of the four "Tier 1" greenhouse gas emissions scenarios known as Shared Socioeconomic Pathways (SSPs) [@RN1451; @RN1452]. The CMIP6 GCM runs were developed in support of the Sixth Assessment Report of the Intergovernmental Panel on Climate Change IPCC AR6. This dataset includes downscaled projections from the [35 models and scenarios](https://pcmdi.llnl.gov/CMIP6/) for which daily scenarios were produced and distributed under CMIP6. 

The Bias-Correction Spatial Disaggregation BCSD method used in generating the [NEX-GDDP-CMIP6 dataset](https://www.nccs.nasa.gov/services/data-collections/land-based-products/nex-gddp-cmip6) is a statistical downscaling algorithm specifically developed to address the current limitations of the global GCM outputs [@RN1282;@RN736;@RN1393;@RN1283]. The NEX-GDPP-CMIP6 climate projections is downscaled at a spatial resolution of 0.25 degrees x 0.25 degrees (approximately 25 km x 25 km). The NEX_GDPP_CMIP6 downscales the NEX-GDPP-CMIP6 data to grid points of 0.1 degrees x 0.1 degrees following nearest point methods described by @RN1207.

## Basic use

Let's use the example watersheds that we introduced with `GPMswat` and `GPMpolyCentroid`. Please visit _NASAaccess_ [GPM](.\GPM.html) functions for more information.

```{r, eval=FALSE, echo=TRUE}
library(NASAaccess)

NEX_GDPP_CMIP6(Dir = "./NEX_GDPP_CMIP6/", 
               watershed = shape_path,
              dem_path,  
              start = "2060-12-1", 
              end = "2060-12-3",
              model = 'IPSL-CM6A-LR', 
              type = 'tas', 
              slice = 'ssp245')
```

Let's examine the air temperature station file

```{r}
NEX_GDPP_CMIP6.temperature.Master <- system.file('extdata/NEX_GDPP_CMIP6',
                                         'tasGrid_Master.txt', 
                                         package = 'NASAaccess')

NEX_GDPP_CMIP6.table<-read.csv(NEX_GDPP_CMIP6.temperature.Master)

head(NEX_GDPP_CMIP6.table)

dim(NEX_GDPP_CMIP6.table)
```


## Built with

```{r}
sessionInfo()


```


## References
